On the Mathematical Formulation of the Angiogenesis Process in Health and Disease

### 2024-09-10 ReviewPaper_Angiogenesis/main.tex
simulation added

	\section{Adaptive Hagen-Poiseuille flows on graphs}
	In this part, I will summarize the methods and the results of the paper \cite{Almeida2022}, and the corresponding theses came out of the same research group that contains more details \cite{Valente2023,Almeida2023}. The Hagen-Poiseuille equation is for the steady-state laminar flow of incompressible, Newtonian fluid through a \textbf{long} channel \footnote{Note that the blood is not a Newtonian fluid and experiences shear thickening (or shear thinning) under different shear rates. But there are certain limits for the shear rate over which we can assume that the blood is a Newtonian fluid. For instance, at a very low shear stress, the viscosity of blood increases a lot at lower tube diameters as the RBC can not squeeze through narrow channels.}. The Hagen-Poiseuille flow is given by
	\[ Q_{ij} = \frac{\pi r_{ij}^4}{8\eta L_{ij}}(p_i - p_j). \]
	For a derivation of this equation from the first principles see \cite{tecscience2020}. The choices for adaptation rule can be due to \cite{Tero2010} where Tero et al propose adaptation rule of the type
	\[ \frac{d D_{ij}}{dt} = f(|Q_{ij}|) - D_{ij}, \]
	where the first term is for the increase of conductance of the tube due to flow and the second term is the baseline shrinkage of the vessel segment. As suggested by Tero et al, one possible choice for $ f(|Q_{ij}|) $ is 
	\[ f(|Q_{ij}|) = \frac{|Q_{ij}|^\gamma}{1+|Q_{ij}|^\gamma} \]
	Similar to what we derived at \autoref{Hu_Work}, the total volume of the network is
	\[ V = \beta \sum_{(i,j)\in E} \sqrt{D_{ij}}L_{ij}. \]
	Note that the variable $ D_{ij} $ represents the conductivity here whereas in \autoref{Hu_Work} it is used for the diameter. By using the chain rule we can write
	\[ \frac{d V}{dt} \propto \sum_{(i,j)\in E} \frac{L_{ij}}{2\sqrt{D_{ij}}} \frac{dD_{ij}}{dt} \]
	where we choose $ d D_{ij}/dt $ according to the ODE governing the time evolution of $ D_{ij} $. Note that not every choice of the RHS of this ODE will guarantee that $ dV/dt = 0 $, and thus the volume is conserved. \cite{Almeida2023} claims that since the choice of $ f $ by \cite{Tero2010}, does not conserve the volume, thus it violates the fact that the fluid is incompressible. I do not think this is true. The volume of the network can shrink, and this has nothing to do with the compression of the fluid inside. Smaller network carries less fluid.

	\section{Work by Karen Alim}

>>>>

	\section{Adaptive Hagen-Poiseuille flows on graphs}
	In this part, I will summarize the methods and the results of the paper \cite{Almeida2022}, and the corresponding theses came out of the same research group that contains more details \cite{Valente2023,Almeida2023}. The Hagen-Poiseuille equation is for the steady-state laminar flow of incompressible, Newtonian fluid through a \textbf{long} channel \footnote{Note that the blood is not a Newtonian fluid and experiences shear thickening (or shear thinning) under different shear rates. But there are certain limits for the shear rate over which we can assume that the blood is a Newtonian fluid. For instance, at a very low shear stress, the viscosity of blood increases a lot at lower tube diameters as the RBC can not squeeze through narrow channels.}. The Hagen-Poiseuille flow is given by
	\[ Q_{ij} = \frac{\pi r_{ij}^4}{8\eta L_{ij}}(p_i - p_j) = \frac{D_{ij}}{L_{ij}}(p_i - p_j). \]
	For a derivation of this equation from the first principles see \cite{tecscience2020}. The choices for adaptation rule can be due to \cite{Tero2010} where Tero et al propose adaptation rule of the type
	\[ \frac{d D_{ij}}{dt} = f(|Q_{ij}|) - D_{ij}, \]
	where the first term is for the increase of conductance of the tube due to flow and the second term is the baseline shrinkage of the vessel segment. As suggested by Tero et al, one possible choice for $ f(|Q_{ij}|) $ is 
	\[ f(|Q_{ij}|) = \frac{|Q_{ij}|^\gamma}{1+|Q_{ij}|^\gamma} \]
	Similar to what we derived at \autoref{Hu_Work}, the total volume of the network is
	\[ V = \beta \sum_{(i,j)\in E} \sqrt{D_{ij}}L_{ij}. \]
	Note that the variable $ D_{ij} $ represents the conductivity here whereas in \autoref{Hu_Work} it is used for the diameter. By using the chain rule we can write
	\[ \frac{d V}{dt} \propto \sum_{(i,j)\in E} \frac{L_{ij}}{2\sqrt{D_{ij}}} \frac{dD_{ij}}{dt} \]
	where we choose $ d D_{ij}/dt $ according to the ODE governing the time evolution of $ D_{ij} $. Note that not every choice of the RHS of this ODE will guarantee that $ dV/dt = 0 $, and thus the volume is conserved. \cite{Almeida2023} claims that since the choice of $ f $ by \cite{Tero2010}, does not conserve the volume, thus it violates the fact that the fluid is incompressible. I do not think this is true. The volume of the network can shrink, and this has nothing to do with the compression of the fluid inside. Smaller network carries less fluid.

	The core of the model discussed in \cite{Almeida2023} is an undirected, planar, and connected graph $ G(V,E) $ that is embedded in Euclidean space. This models is originally developed for the 2D vein system of leaf. But in the case of blood flow, it should be a 3D graph. Also, they have assumed that the edges are straight. The convention they have used here is that $ Q_{ij}>0 $ means a flow from node $ i $ to $ j $, and vice versa. The sources and the sinks are located at fixed nodes. In the case of blood networks, the sinks are the cells distributed through the domain. We need to figure out some way to model this effectively as sinks on the nodes. The conservation law reads
	\[ \sum_{j\in\mathcal{N}(i)}Q_{ij} = s_i. \]
	\textbf{Local vs. Global Mechanisms:} The authors argue that the decision to conserve the volume through the dynamics is considered a global mechanism. That is because if some of the vessels shrink due to the lack of flow, then some other vessel should necessarily expand to keep the total volume conserved. They argue that they mechanism proposed by \cite{Tero2010} is a local dynamic.

	Since the adaptation of each vessel segment is due to the change it its cross section (radius) and since $ r_{ij} \propto \sqrt{D_{ij}} $, thus we write the ODE for $ \sqrt{D_{ij}} $. One initial suggestion is
	\[ \frac{d}{dt} \sqrt{D_{ij}} = f(Q) - \mu \sqrt{D_{ij}},\]
	where $ f $ is an unknown function with $ f(0) = 0 $. Then we will need to make sure that this adaptation rule conserves the total volume of the network.

	To find the pressures on the nodes, given the values of the sources/sinks on the nodes, as well as the conductance of the links, for each node $ i $ we need to solve
	\[ \sum_{j\in\mathcal{N}(i)}Q_{ij} = \sum_{j\in\mathcal{N}(i)}\frac{C_{ij}}{L_{ij}}(p_i - p_j) = s_i, \]
	which is a kind of Poisson equation on the graph. This can be formulated as
	\[ L \mathbf{P} = \mathbf{q}, \]
	where $ L = D - A $ is the Laplacian matrix of the graph, where $ D $ is the degree matrix and $ A $ is the adjacency matrix.

	\newpage
	\section{Work by Karen Alim}
